---
title: "I5 Monthly Renewal Rate Report"
output:
  html_document:
    df_print: paged
---

Report created on: `r format(Sys.Date(), "%b %d, %Y")`

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(kableExtra)
library(lubridate)
#library(readxl)
```

``` {r data_wrangle}
dat2 <- read_csv("DATA\\newest_renewal_data.csv") %>%
  rename(member_id = `Constituent ID`,
         joined_date = `Current Joined On Date`,
         renewal_date = `History All Renewed On Dates`,
         curr_expiration_date = `Current Expires On`,
         curr_member_category = `Current Category`,
         member_category = `History Membership Category`,
         expiration_date = `History Expires On`) %>%
  mutate(renewal_date = as.Date(renewal_date, format = "%m/%d/%Y"),
         curr_expiration_date = as.Date(curr_expiration_date, format = "%m/%d/%Y"),
         expiration_date = as.Date(expiration_date, format = "%m/%d/%Y"),
         member_category = as.factor(if_else(is.na(`Gift Reference`), member_category, "CEAP")),
         # member_category = as.factor(member_category),
         joined_date = as.Date(joined_date, format = "%m/%d/%Y")) %>%
  select(member_id, joined_date, renewal_date, curr_expiration_date, expiration_date, member_category) %>%
  arrange(member_id, expiration_date) %>%
  distinct() # the query change that added the gift recipient column led to some duplicates in the report with the history columns, goes away once you select distinct on just the columns we need

cutoff_date <- floor_date(Sys.Date() - months(0), "month") - 1

renewals2<-dat2 %>%
  group_by(member_id) %>%
  mutate(next_renewal = lead(renewal_date, order_by=member_id),
         future_expir = ifelse(expiration_date > Sys.Date(),1,0),
         member_is_numeric = ifelse(!grepl("\\D", member_id),1,0),
         CountInst = n(), 
         membership_year = row_number(),
         renewal_timing = as.numeric(difftime(next_renewal, expiration_date , units = "days")),
         renewed = if_else(is.na(renewal_timing) | renewal_timing > 183, 0, 1)) %>%
  filter(member_is_numeric==1,
         !is.na(member_id),
         !member_category %in% c("Corporate", "SED"),
         expiration_date >= '2018-10-31' & expiration_date <= cutoff_date) %>%
  select(-member_is_numeric) %>%
  arrange(member_id, renewal_date)
```

### Basic renewal rates

A membership purchase is considered a renewal if the transaction occurs within 6 months after the end of a membership expiration. The table below shows the total renewal rate, regardless of the renewal timing, for expiring memberships by type.

The rates are calculated only on memberships that expired at least 7 months ago so that the full rate is known. The data starts with memberships expiring in Oct 2018, the year after the membership structure changed.

``` {r overall_rate, echo=FALSE, message=FALSE}
full_renewals <- renewals2 %>% 
  filter(expiration_date <= floor_date(Sys.Date() - months(6), "month") - 1)

basic_renewal_rate <- full_renewals %>%
  group_by(member_category) %>%
  summarise(`Renewal Rate`=round(mean(renewed),3)*100, `Member Count`=n(), `Last Exp. Month` = format(max(expiration_date), "%b-%Y")) %>%
  ungroup() %>%
  rename(`Member Category` = member_category)

knitr::kable(basic_renewal_rate) %>%
   kable_styling(bootstrap_options = "striped",
                 full_width = F)

overall_renewal_rate <- round(mean(full_renewals$renewed), 3) * 100
```

The renewal rate across memberships is `r overall_renewal_rate`%.

### Detailed Renewal Rates

The table below shows the renewal rate by month, broken out by the timing of the renewal:

* **On Time:** Renewed by expiration date
* **Less than 1 Month:** Within 31 days of expiration
* **Late (w/in 6 mos):** Within 183 days of expiration
* **Overall rate:** Sum across the categories above

This table is also available in a CSV file that can be opened as an excel spreadsheet.

``` {r rate_by_mth}
renewal_by_mth <- renewals2 %>% 
  mutate(renew_type = case_when(is.na(renewal_timing) ~ "Never Renewed",
                               renewal_timing <= 0 ~ "On Time",
                               renewal_timing <= 31 ~ "Less than 1 Month",
                               renewal_timing <=183 ~ "Late (w/in 6 mos)",
                               renewal_timing > 183  ~ "Never Renewed",
                               TRUE ~ "NonCategory"))


renew_dtl <- renewal_by_mth %>%
  group_by(member_category, expiration_date, renew_type) %>%
  summarise(occur = n()) %>% 
  ungroup() %>%
  # remove periods that are not complete
  mutate(occur = case_when(
    expiration_date > Sys.Date() ~ NA_integer_,
    expiration_date > (Sys.Date() - 31) & renew_type == "Less than 1 Month" ~ NA_integer_,
    expiration_date > (Sys.Date() - 183) & renew_type == "Late (w/in 6 mos)" ~ NA_integer_,
    TRUE ~ occur
  ))

mthly_total <- renewal_by_mth %>%
  group_by(member_category, expiration_date) %>%
  summarise(total = n()) %>% 
  ungroup()

mthly_renewal_rate_by_memcat <- renew_dtl %>%
  left_join(mthly_total) %>%
  mutate(`% Renewed` = occur / total)

mthly_renewal_rate_by_memcat_wide <- mthly_renewal_rate_by_memcat %>%
  select(member_category, expiration_date, total, renew_type, `% Renewed`) %>%
  mutate(`Exp. Date` = format(expiration_date, "%b-%Y"),
         `Members` = as.character(total)) %>%
  spread(renew_type, `% Renewed`) %>%
  mutate(`Overall Rate` = `On Time` + `Less than 1 Month` + `Late (w/in 6 mos)`) %>%
  select(member_category, `Exp. Date`, expiration_date, `Members`, `On Time`, `Less than 1 Month`, `Late (w/in 6 mos)`, `Overall Rate`) %>%
  rename(`Member Category` = member_category)


# apply % formatting
vars <- c("On Time", "Less than 1 Month", "Late (w/in 6 mos)", "Overall Rate")
mthly_renewal_rate_by_memcat_wide[vars] <- lapply(mthly_renewal_rate_by_memcat_wide[vars], scales::percent)

  # remove the NA % and the previous month's Less than 1 Month
mthly_renewal_rate_by_memcat_wide <- mthly_renewal_rate_by_memcat_wide %>%
  mutate(#`Less than 1 Month` = if_else(Sys.Date() - `Exp. Date` < 31, "", `Less than 1 Month`),
         `Less than 1 Month` = if_else(`Less than 1 Month` == "NA%", "", `Less than 1 Month`),
         `Late (w/in 6 mos)` = if_else(`Late (w/in 6 mos)` == "NA%", "", `Late (w/in 6 mos)`),
         `Overall Rate` = if_else(`Overall Rate` == "NA%", "", `Overall Rate`)) %>%
  arrange(`Member Category`, desc(expiration_date)) %>%
  select(-expiration_date)

knitr::kable(mthly_renewal_rate_by_memcat_wide) %>%
   kable_styling(bootstrap_options = "striped", full_width = T, fixed_thead = T) %>%
  scroll_box(width = "100%", height = "500px")

report_path <- paste0("OUTPUT\\detailed_renewal_rate_", Sys.Date(), ".csv")

write_csv(mthly_renewal_rate_by_memcat_wide, report_path)

```