---
title: "Understanding Renewal Rates"
# output: html_notebook
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
Serving as the initial analysis of the renewal rates at Impression 5 from 2014-2018

```{r setup,include=FALSE,echo=FALSE,message=FALSE}
library(tidyverse)
library(lubridate)
library(ggthemes)
library(here)
```


``` {r data_wrangle,include=FALSE,echo=FALSE,message=FALSE}
# dat<-read_csv("DATA\\member_renewals_2014-2018.csv") %>%
IMP <- read.csv(here("Impression5-master", "DATA", "member_renewals_2014-2018.csv"), stringsAsFactors = FALSE) %>% 
  rename(member_id = `Constituent.ID`,
         renewal_date = `History.All.Renewed.On.Dates`,
         expiration_date = `History.Expires.On`,
         member_category = `History.Membership.Category`) %>%
  mutate(renewal_date = as.Date(renewal_date,format="%d/%m/%Y"),
         expiration_date = as.Date(expiration_date,format="%d/%m/%Y"),
         member_category = as.factor(member_category)) %>%
  mutate(member_is_numeric = ifelse(!grepl("\\D", member_id),1,0)) %>%
  filter(member_is_numeric==1,
         !is.na(member_id), !member_id  =="",
         !member_category=="Corporate") %>% #drop non-numeric member id, NA member id, and first year
  arrange(member_id, expiration_date) %>%
  group_by(member_id) %>%
  mutate(CountInst = n(), RewCount = row_number()) %>%
  mutate(next_renewal = lead(renewal_date, order_by=member_id),
         future_expir = ifelse(is.na(next_renewal) & expiration_date > (Sys.Date()-30),1,0)) %>%
  mutate(RenewComp = difftime(next_renewal, expiration_date , units = "days"),
         RenewComp = as.numeric(RenewComp)) %>%
  mutate(RenewMonth = month(expiration_date)) %>%
  ungroup()


```

First step was to see when in general the expiration dates are and what the common member categories are.

From this view, it looks like expiration dates (~1 year after signing up) are more common in December through April and the most common member category being "Family".

``` {r analysis}
IMP %>% select(member_id, RenewMonth, member_category) %>% unique() %>% group_by(RenewMonth,member_category) %>% summarise(CountMo = n()) %>% 
  ggplot(aes(RenewMonth, CountMo, fill=member_category))+
  geom_bar(stat="identity")+
  scale_x_continuous(breaks= c(1:12))+
  theme_bw()+
  ggtitle("Month of Expiration Date by Member Category [1: Jan, 12: Dec]")

```

There were three main renewal types:  
1. Member renewed on time (renewed in less than 30 days from expiration date)   
2. Member renewed late (renewed between 30 days and 6 months after expiration date)   
3. Member didn't renew (renewed after 6 months or never renewed)  

For the sake of the analysis another category was added ("NeverRenew") to captured instances where a member signed up and didn't renew after that.

```{r Trend Renewal Types, fig.width= 10}
remFuture <- IMP %>% 
  filter(future_expir == 0) %>% 
  mutate(RenewType = case_when((RewCount == 1 & is.na(RenewComp)) ~ "NeverRenew",
                               RenewComp <= 30 ~ "OnTime",
                               (RenewComp > 30 & RenewComp <=180) ~ "Late",
                               RenewComp > 180 | (RewCount > 1 & is.na(RenewComp))  ~ "RenewTooLate_Didnt",
                               TRUE ~ "NonCategory"))

remFuture %>% group_by(RenewType) %>% summarise(Occur = n()) %>% 
  ungroup()


remFuture %>% mutate(RenewMonth = month(expiration_date)) %>% select(member_id, RenewMonth, RenewType) %>% unique() %>% 
  group_by(RenewMonth,RenewType) %>% summarise(CountMo = n()) %>% 
  ggplot(aes(RenewMonth, CountMo, fill=RenewType))+
  geom_bar(stat="identity")+
  scale_x_continuous(breaks= c(1:12))+
  theme_bw()+
  ggtitle("Renewal type by month (all years)")

remFuture %>% mutate(RenewMonth = month(expiration_date), RenewYr = year(expiration_date)) %>% select(member_id, RenewMonth, RenewYr, RenewType) %>% unique() %>% 
  group_by(RenewMonth,RenewYr, RenewType) %>% summarise(CountMo = n()) %>% 
  filter(RenewYr < 2019) %>% 
  ggplot(aes(RenewMonth, CountMo, fill=RenewType))+
  geom_bar(stat="identity")+
  scale_x_continuous(breaks= c(1:12))+
  facet_grid(~RenewYr)+
  theme_bw()+
  ggtitle("Renewal type by month (by year)")

```

```{r Percent Renewals by Month , fig.width= 10}
stackPer <- remFuture %>% mutate(RenewMonth = month(expiration_date)) %>% select(member_id, RenewMonth, RenewType) %>% unique() %>% 
  group_by(RenewMonth,RenewType) %>% summarise(CountMo = n()) %>% 
  mutate(ratio=scales::percent(CountMo/sum(CountMo))) %>%  ungroup()

  ggplot(data= stackPer, aes(RenewMonth, CountMo, fill=RenewType))+
  geom_bar(position = "fill",stat="identity")+
  geom_text(data=stackPer, aes(y=CountMo,label=ratio),
            position=position_fill(vjust=0.5))+
  scale_x_continuous(breaks= c(1:12))+
  scale_y_continuous(labels = scales::percent)+
  theme_bw()+
  ggtitle("Percent of Renewal type by month (all years)")  
```

```{r Percent Renewals by Month by Year, fig.width= 12, fig.height= 12}
stackPerYR <- remFuture %>% mutate(RenewMonth = month(expiration_date), RenewYr = year(expiration_date)) %>% select(member_id, RenewMonth, RenewYr,RenewType) %>% unique() %>% 
  group_by(RenewMonth, RenewYr,RenewType) %>% summarise(CountMo = n()) %>% 
  mutate(ratio=scales::percent(CountMo/sum(CountMo))) %>%  filter(RenewYr < 2019) %>%  ungroup()

ggplot(data= stackPerYR, aes(RenewMonth, CountMo, fill=RenewType))+
  geom_bar(position = "fill",stat="identity")+
  geom_text(data=stackPerYR, aes(y=CountMo,label=ratio),
            position=position_fill(vjust=0.5))+
  scale_x_continuous(breaks= c(1:12))+
  scale_y_continuous(labels = scales::percent)+
  theme_bw()+
  facet_grid(RenewYr~.)+
  ggtitle("Percent of Renewal type by month (by year)")  
```

